% -*- mode: latex; mode: auto-fill; coding: utf-8; -*-

\chapter*{Acknowledgements}
We greatly appreciate the help of all who contributed to this thesis,
especially the following people whom we would like to give special
thanks to: \\ 

Ole Østerby, Associate Professor, for supervision, guidance, and
for always leaving his door open, both literary and figuratively. We
are grateful for the time he spent carefully explaining theories
to us and for his comprehensive reviews. 
%
Thomas Sangild Sørensen, Associate Professor, for supervision,
especially within the field
of simulated surgeries and for his inspiring talks, ideas,
references, reviews, and guidance throughout this entire thesis.
%
Karsten Østergaard Noe, Ph.D. student, for sharing his knowledge and
a prototype solver for structural analysis. \\

Jesper Mosegaard, Ph.D., and Peter Trier, both from The Alexandra
Institute, for introducing the problem domain, for their cooperation and
for establishing collaboration with people from the Aarhus School of
Dentistry, Aarhus University. \\

Søren Schou, Professor and dr.~odont., at Aarhus School of Dentistry, Aarhus
University, for his enthusiasm and the time he spent on carefully
explaining how dental surgical procedures are performed. We are
thankful for all the material he provided including books, videos,
dental instruments, wisdom teeth and for granting us permission to
use the dental pictures included in this thesis (marked by \textdagger). \\

Lau Brix, Ph.D. student, for guidance concerning hardware
configuration for a high performance computer suitable for parallel
execution
%
and the Siemens Foundation (Siemens fonden) for funding this
hardware.
%
Christian Esbo Agergaard, 3D artist, for creating a three-dimensional
surface mesh of a wisdom tooth, which is suitable for our simulation
scenarios.
%
The OpenEngine community for providing the portable framework used as
the foundation for the simulator implementation. \\

Thanks are also do to the librarians, technicians, and administrative
staff at Aarhus University for their help and effectiveness.
